\documentstyle[12pt]{article}

\setlength{\topmargin}{-.7in}
\setlength{\leftmargin}{-.7in}
\setlength{\textheight}{9in}
\setlength{\oddsidemargin}{0in}
\setlength{\textwidth}{6.25in}

\begin{document}

\title{Design Brief: Funderwater Corporation Music Project}
\maketitle
Client: Funderwater Corporation

Designer: Cameron Carroll (Engr 100 Design Team)

Design Statement: Design a device that a diver can easily see and manipulate, enabling them to listen to music.

\vspace{.2in}
\hrule
\vspace{.2in}

{\large Constraints:}
\begin{itemize}
\item Allow a diver to listen to music underwater and be able to change to song. (Measured by underwater test.)
\item Must be operable through a variety of diving gloves. (Measured by underwater glove test.)
\item Must cost less than \$20 to manufacture each unit. 
\item Must be reasonably compact \& light. (Measured by diver panel.)
\end{itemize}

\vspace{.3in}

{\large Objectives:}
\begin{itemize}
\item Materials should be available \& in a competitive market. (Measured by supplier count.)
\item Buttons should have 'feedback' to allow for 'blind' use. (Measured by blind button test.)
\item Should follow interference \& safety standards. (Measured by internal audit.)
\item Should have a colorful / tropical feel. (Measured by consumer panel.)
\item Should be manufacturable. (Measured by consulting manufacturer.)
\end{itemize}

\vspace{.3in}

{\large Deliverables:}
\begin{itemize}
\item Design due October 13, 2011
\end{itemize}
\end{document}

