\documentclass[11pt]{article}

\setlength{\topmargin}{-1.5in}
\setlength{\leftmargin}{-.7in}
\setlength{\textheight}{9in}
\setlength{\oddsidemargin}{0in}
\setlength{\textwidth}{6.25in}

\newcommand{\squishlist}{
 \begin{list}{$\bullet$}
  { \setlength{\itemsep}{0pt}
     \setlength{\parsep}{3pt}
     \setlength{\topsep}{3pt}
     \setlength{\partopsep}{0pt}
     \setlength{\leftmargin}{1.5em}
     \setlength{\labelwidth}{1em}
     \setlength{\labelsep}{0.5em} } }

\newcommand{\squishend}{
  \end{list}  }

\usepackage{times}

\begin{document}

\title{Design Brief: Barkley Concepts, Inc: Water Conservation Project}
\maketitle
\paragraph{Client:} Barkley Concepts, Inc.

\paragraph{Designer: } Engineering 100 Design - Team Super 8-Track
\paragraph{Design Team:}
\squishlist
\item Cameron Carroll \\[-0.4cm]
\item Daniel Sharon \\[-0.4cm]
\item Jonathan Kohlenberg \\[-0.4cm]
\item Mark Richards \\[-0.4cm]
\squishend

\hrule

\paragraph{Design Statement: } Design a product that helps conserve water and pays for itself within five years. \\[-0.4cm]

\paragraph{Constraints: }
\squishlist
\item Reduces Water Consumption, measured by a gallon comparison against unmodified system. \\[-0.4cm]
\item Pays for itself within 5 years, measured by comparison against 0.5 cent per gallon metric \& 127,000 gallons used every year, on average. \\[-0.4cm]
\item Meets applicable safety/health codes and standards, as measured by internal audit. \\[-0.4cm]
\item Product costs no more than \$250 \\[-0.4cm]
\squishend


\paragraph{Objectives: }
\squishlist
\item As easy to use as possible, measured by consumer panel. \\[-0.4cm]
\item As cheap as possible, measured in price difference in USD. \\[-0.4cm]
\item As safe as possible, measured by test installation. \\[-0.4cm]
\item Materials should be highly available, measured by number of available sources.\\[-0.4cm]
\item Cheap to manufacture, measured by difference in manufacturing cost. \\[-0.4cm]
\item Minimize installation cost, measured by difference in average installation cost. \\[-0.4cm]
\squishend



\paragraph{Deliverables: }
\squishlist
\item Design presentation due December 6, 2011 \\[-0.4cm]
\item Final product design due June 6, 2011 \\[-0.4cm]
\squishend
\end{document}

