\documentclass[12pt,letterpaper]{article}

\usepackage{ifpdf}
\usepackage{mla}

\begin{document}
\begin{mla}{Cameron}{Carroll}{Prof. Mood}{English 124}{\today}{Worldview Paper: Reflection}

The writing process was somewhat slow to start for me. There are so many topics I wanted to consider on this topic...For example,  I didn't really find room to start talking about cultural legacies. I also found a new environment for writing and typesetting in LaTeX and am thoroughly pleased. MLA formatting is ridiculously easy, and I find it easier to focus on the text as opposed to Word. As always, the first few passes through were just to get the ideas down and recorded; Later iterations were focused on polishing language, pace and order. 

The peer review workshop was fairly helpful; My peers caught a few grammatical errors and helped to pinpoint where the paper stopped being cohesive. My biggest problem going into it was that all of my paragraphs had different topics and directions and almost nothing tied together. A discussion with them as well as reading their papers gave me a little direction for cleaning it up. I didn't formally record their names, but I worked with Kandee, Cissy and the gentleman whose name I don't remember, but hails from Bakersfield.

My paper has good language but poor cohesion. A lot of the ideas are kind of a smorgasborg of brief physiological research (I mean, ultra brief) and minor sociological knowledge/research. Essentially, I feel like I sound much more informed and sure of myself than I actually am.

This paper forced me to think about myself a lot more than usual and encouraged introspection. It's always fascinating to ponder where our ideas come from and where they're going, but can we ever know definitively? This paper helped me as a writer because I hadn't really been looking between the lines with the language and ideas: Sure, the basics are there, but what are they really saying and how is the way it is said influence the idea? I wasn't looking for the roots, and for the pace and the length of sentences and how everything flows together. I wish I had more time because this paper isn't done. I wouldn't feel good about its life, should it fall into a puddle of toxic waste and spring legs... It sounds good, and the ideas are there, but it isn't beautiful. Writing is like that, for me: Whether it's code, or prose, or poetry or math... conveyance is only half of the purpose. The other half is simply the beauty of the expression.


\end{mla}
\end{document}