\documentclass[13pt]{article}


\usepackage{fourier}
\usepackage{sectsty} % Allows customizing section commands
\allsectionsfont{\centering \normalfont\scshape} % Make all sections centered, the default font and small caps
\usepackage{graphicx}
\usepackage{mhchem}
\usepackage{tabularx}
\usepackage{pbox}
\usepackage{listings}
\lstset{language=Matlab}
\usepackage{caption}
\usepackage{color}
\usepackage{xcolor}
\usepackage{fullpage}
\linespread{2}

\newcommand{\HRule}{\rule{\linewidth}{0.5mm}}


\begin{document}

\input{./movie2_title.tex}
\pagebreak

\paragraph{} In `Lincoln,' we follow the efforts of the Union and the man himself toward ratification of the thirteenth amendment, where he fights through myriad battles in the trenches of politics while his soldiers debate in the field.

\paragraph{(Question 1)} There are a couple moods of notes portrayed in this film; There's the juxtaposition between the politicians and the soldiers, and there's also the black man enslaved, freed, and born free. \\
\paragraph{} The former comparison shows vigorous argumentation by the aristocrats, lawyers and politicians. They are essential, of course, to strategize and drive the legal and political side of the battle, but their energy is in stark comparison to what appear to be generally tired and worn soldiers.
\paragraph{} The latter shows the differences between free blacks and those enslaved; Free blacks are shown as having a feeling of superiority toward their enslaved or freed counterparts. They're not quite free white men landowners, but they seem to want to distance themselves from the institution. For those freed, the wounds are still quite fresh.


\paragraph{(Question 2)} Sort of building on the previous idea, I found similarities between the film's portrayal and today in both the politics and the war. There's a clear line between those in the war and those at home; At this very moment we have American soldiers at various places off in the rest of the world, in conflict, and we're not particularly concerned. Likewise, in this film life goes on everywhere the war isn't -- In particular, those in power are making decisions for the entire populace with an obvious removal from their concerns. Sure, someone has to fight and other must make decisions and laws, but the differences are clear between those arguing at our capital and those at gunpoint.

\end{document}
