\documentclass{article}
\begin{document}
  
  \begin{center}
    {\small{} Early American History Lecture Notes} \\[0.6cm]
    {\small{} Cameron Carroll -- August 28, 2013} \\[0.6cm]
    {\small{} Lecturer: Judy Campbell, Cuyamaca College}\\[1cm]
    {\small{} Early English Colonization I}\\[1cm]
  \end{center}
  
  \tableofcontents
  \newpage

  \section{Early Colonization by England}

    \begin{itemize}
      \item She came to America for gold, emulating Spain, but didn't find any. Instead decided tobacco was worth pursuing. 
      \item `Youthful Lusties' went to New World looking for gold. 
      \item Spain's been getting rich for the last like hundred years, England couldn't fight. England hires `sea dogs' (pirates). These would raid Spanish galleons to get money to send people to New World. Eventually leads to sinking of Spanish armada in ~1500s, and England becomes power of Europe.
    \end{itemize}

    \subsection{Reasons for Colonization}
      \begin{itemize}
        \item War of Roses; English were fighting among themselves trying figuring out who will rule.
        \item Just starting the industrial revolution at this point. Figured the New World would make good dependents upon English goods, so they can create markets for capitalism.
        \item Succinctly, England is coming to settle; Investors needed a place to put their capital.
        \item Hakluyt's Discourse on Western Planting:
          \begin{itemize}
            \item Was a writing encouraging the government to adopt a policy of maritime trade and colonial development.
            \item Impels the English to plant settlements funded by joint stock companies.
          \end{itemize}
        \item Joint Stock Companies:
          \begin{itemize}
            \item Were rich people coming together, pooling stock/money in order to send people to the New World. Shipbuilding/manning was expensive and supplies for ships/people were expensive.
            \item Most settlements in the beginning were through joint stock companies... nobody could come by themselves. Became known as venture capitalists.
          \end{itemize}
        \item Mercantilism: A philosophy where you create favorable balance of trade. We send manufactured goods somewhere and get back natural resources. England went to America to fulfill this philosophy.
        \item England was pretty much full, land was depleted, and it had been settled for a long time. She had a population boom with nowhere to go, and so the New World with its ample farmland became an attractive option. The government liked the idea of taxes on goods in \& out.
      \end{itemize}

    \subsection{Movers \& Shakers}
      \begin{itemize}
        \item Luther: Wrote 95 things he thought the church was doing wrong.
        \item Calvin: Instrumental in new england colonies... comes up with predestination where some are destined for heaven, others hell and you're stuck. This idea gets carried to new england, especially Massachusetts where only church members can be in government and vote. (Because if you're not a church member you're doomed.)
        \item Henry VIII: Got divorced like 5 times. The pope told him that's not okay, Henry disregarded him. Goes through 6 wives trying to get a male heir. Starts his own church, church of England (a protestant church) because the pope is on his ass. Repressed everything except Church of England (so all the pilgrims dipped out)
        \item Elizabeth I: Wasn't a bitch. Didn't persecute quite as many people. 
      \end{itemize}

    

    \subsection{First Settlements}
      \subsubsection{Roanoke Island}
        \begin{itemize}
          \item The very first English settlement in North America, Roanoke Island lies off the coast of North Carolina.
          \item English were short money so they gave land to people in the New World. Raleigh gets given Roanoke, goes over with a grip of people but they don't last.
          \item In later 1500s, Raleigh tries again, takes ~90 colonists and try to settle same island again. While there sets up John White as leader of island. 
          \item John White has a granddaughter born. Virginia Dare is first child of European descent born in what is now America.
          \item White discovers they aren't gonna survive with their supplies. White goes bck to England to get supplies, gets delayed like three freaking whole years. Comes back to Roanoke and finds nobody, nothing, no evidence. (Although supposedly they just went to the village down the road...)
        \end{itemize}

      \subsubsection{Jamestown, Virginia}
        \begin{itemize}
          \item 1607, a land grant was given to London Company, a joint stock company, to create an American settlement for England.
          \item Jamestown was first permanent English colony.
          \item 40\% of persons coming here were out-of-work nobles, jewelers, etc looking for easy money. They had no idea how to survive in the New World.
          \item Jamestown was created: A little fort next to a malarial swamp.
          \item They tried to have decent relations with natives, but were too greedy.
          \item First colonists lost money, so company sent more ships and colonists but still only sent `useless' men.
          \item Around 1600 came the `starving time.' They ate their horses, shoes, and each other.
          \item They sent John Smith to try to fix things, who took Lord DeLaWare with him. These two guys started draconian measures to get Jamestown into shape:
            \begin{itemize}
              \item You steal corn $\rightarrow$ You lose your hand.
              \item You run away to the Indians $\rightarrow$ You get killed.
              \item You don't work $\rightarrow$ You don't eat.
            \end{itemize}
            But that's basically what the town needed after only surviving off of native charity.
          \item Ultimately, Jamestown survived with help of the Powhatan tribe \& Pocahontas in particular:
            \begin{itemize}
              \item Pocahontas is infatuated with John Smith, warns Jamestown of impending Indian attack but gets left behind by him. (I mean come on she was like twelve.)
              \item Then meets Rolfe, gets taken to England, becomes adored by the court of England and has a son.
              \item Then wants to go back to Jamestown/Virginia, but dies the night before she sailed.
              \item Her son later becomes governor of Virginia.
            \end{itemize}
          \item By 1612, a milder form of tobacco was introduced and becomes popular:
            \begin{itemize}
              \item Becomes a money source for Jamestown.
              \item The plant, however, depletes the soil, so Virginia colonists have to keep moving inland to find more land. (Taking Indian land).
              \item Tobacco and rice are very labor intensive, leading to forced labor. The concept of indentured servitude starts: A person whose passage is paid in exchange for ~4-5 years of service. At the end of their service they could go freely. 
            \end{itemize}
          \item In 1676, Bacons Rebellion happens:
            \begin{itemize}
              \item King sent over a dude Berkeley to oversee the colony.
              \item The farmers needed more land, the king said no don't, and Berkeley wouldn't give permission.
              \item But the farmers did it anyway! They killed all the Indians they could find, including the ones that helped them before.
              \item The indians in turn had uprising against colonists.
              \item Soooo the colonists called for a peace treaty... invited ~250 indians to hang out. But they poisoned all the indians.
              \item Then the indians tried to burn down Jamestown, and it was all a mess.
            \end{itemize}
        \end{itemize}

      \subsubsection{Plymouth, Massachusetts}
        \textbf{(On the next episode of Early American History.)}
\end{document}