\documentclass{article}
\begin{document}
  
  \begin{center}
    {\small{} Early American History Lecture Notes} \\[0.6cm]
    {\small{} Cameron Carroll -- September 4, 2013} \\[0.6cm]
    {\small{} Lecturer: Judy Campbell, Cuyamaca College}\\[1cm]
    {\small{} Early English Colonization II}\\[1cm]
  \end{center}
  
  \tableofcontents
  \newpage
  
  \section{English Colonies in North America}
    \subsection{Plymouth}
      \begin{itemize}
        \item Settlers were sent by the Plymouth company, a joint-stock company, and were suppose to go to Jamestown. Instead they're blown to Massachusetts by a storm and decide to go to storm.
        \item Fun fact: They landed partially due to a lack of beer, and set up a brewery almost immediately.
        \item Notable for being actually gracious and thankful to the Indians helping them. This is in sharp contrast with Jamestown. There was no enslavement of Indians; They treated them fairly. 
        \item They had work ethics: everyone worked for food.
        \item First Thanksgiving was a three day party giving thanks to god for making it through the winter. Half of the pilgrims had died but the other half threw a party after winter was over.
      \end{itemize}

      \subsubsection{Separatists}
        \begin{itemize}
          \item The settlers called themselves separatists \ldots They were Protestants who wanted to separate from the Church of England, which had separated from the Catholic church.
          \item They were part of protestant reformation\ldots known as pilgrims but not given that name for another 70 years.
          \item Came over on the Mayflower, which holds 102 people in incredibly small quarters. Also on the boat were `adventurers/strangers'
          \item All under authority of William Bradford. 
        \end{itemize}

        \subsubsection{Mayflower Compact}
          \begin{itemize}
            \item Was a written agreement for conduct among themselves.
            \item Was the first time people had determined a self-government worldwide.
            \item Only a page long.
            \item It acknowledged they were under King and God and made a religious settlement of families, but still determined rules for their own existence.
            \item Only men were allowed to sign it, and no strangers were allowed to sign but still had to obey. 
          \end{itemize}

        \subsubsection{Squanto}
          \begin{itemize}
            \item Was an Indian who had been captured and was taken to Europe. He is taken back to America, escapes back to his tribe.
            \item The Pilgrims then  run into a tribe who has this dude who speaks perfect English. He then acts as liasion between English pilgrims and indians. 
          \end{itemize}

    \subsection{Massachusetts Bay Colony}
      \subsubsection{Colonization}
        \begin{itemize}
          \item Settled around 1630 under John Winthrop.
          \item Gobbles up smaller colony-lets around it (Salem.) 
          \item About 25,000 people by 1640. Undergoes a huge sudden influx of people in the last 20 years.
          \item Consisted of mostly puritans, some calvinists. They tended to be intolerant towards other religions despite escaping their own persecution. Had to belong to church in order to have a voice in government.
          \item In New England, people settle in towns instead of farms. Geography (lack of land) forces families into smaller areas, i.e. towns. Mountains leads to running water leads to factory development. Lots of harbors and inlets lead to shipping. Timber from the abundant forests help economy of Mass. Bay Colony.
          \item In the south, rivers are flat and meandering: Good for agriculture, river trade. There are some ports but the agricultural land leads to our well known labor trends.
          \item Colony had 70,000 Englishmen by 1680.
        \end{itemize}

      \subsubsection{Trading System: Triangle Trade}
        \begin{itemize}
          \item England sends goods to America, which returns raw material.
          \item England is all about sugar at the time... after the Columbian Exchange, Europeans who never had access got used to sugar and molasses.
          \item England sends manufactured goods to Africa, which sends people to America.
          \item Lucky few slaves go to the Colonies, while the unlucky masses go to die in the Carribbean.
          \item African dudes at this time would trade fellow African captives to the west coast and sell them into slavery for material goods. (Like the Indians, they were not in the iron age.) Portugese validated the slavery by christianizing the slaves. 
          \item Slavers usually lost about 1/3 of cargo, kept them barely alive, chained together in too-short quarters for months. (Middle passage) Fun fact: The sharks learned the shipping routes knowing they would get bodies.
        \end{itemize}

  \subsection{New Hampshire}
      \begin{itemize}
        \item Separates from MBC for a little while, establishes plantations but they don't make it.
        \item Gets reabsorbed by MBC for another 50 years.
      \end{itemize}

    \subsection{Rhode Island}
      \begin{itemize}
        \item Settled by Roger Williams and Anne Hutchinson. Both exiled from MBC for being radical -- called `extremes.' Sent down to bottom of MBC to Rhode's Island. 
        \item Had complete religious freedom, civil rights, church \& state separation. Did not require you to belong to church to have a voice in government. Purchased land from Indians.
        \item Anne Hutchinson started women's bible study. Puritans were horrified by the idea that a man would ever be under a woman.
      \end{itemize}

    \subsection{Connecticut}
      \subsubsection{Colonization}
        \begin{itemize}
          \item Thomas Hooker was another guy rejected from the MBC. So he took ~100 people down to New Haven, botom of MBC. 
          \item Some Indians happened to be under assault by their neighbors and get Hooker to defend them in exchange for living on the land.
          \item A bible commonwealth is then established.
        \end{itemize}
      \subsubsection{Fundamental Orders of Connecticut}
        \begin{itemize}
          \item First written constitution in what is now the United States.
          \item Said that social order are to determine their own rule.
          \item Established government \& gave all people a voice. By all people they mean all male white property owners.
        \end{itemize}
      \subsubsection{King Philip's War}
        \begin{itemize}
          \item As people fill up New England area, Indians become displaced and perturbed.
          \item In the 1660s there's King Philip's war. (King Philip is an indian.)
          \item This was a bloody retaliation against settlers moving into the mountains -- Killed ~500 settlers. Colonists retaliated and killed ~3000 Indians.
        \end{itemize}

    \subsection{New York}
      \begin{itemize}
        \item Settled by the Dutch West India Company (joint stock) first. In 1620 or so, Dutch set up a governor: Peter Minuet. He purchased Manhattan Island from the Indians for \$24. He then named the area New Amsterdam.
        \item At this time (early 1600s) English were fighting the Dutch, and English win around 1660-1670. As a result, England is given New Amsterdam. The Duke of York was in the king's favor and was given New Amswerdam... didn't really care as long as it brought some wealth. 
      \end{itemize}

    \subsection{New Jersey and Delaware}
      \begin{itemize}
        \item Swedes, English, Dutch, Quakers and Puritans all in here. Cheap land, nothing exciting.
      \end{itemize}

    \subsection{Maryland}
      \begin{itemize}
        \item 1630's -- Settled by catholics. King gave Maryland to George Calvert, known as a proprietor. King giving land is known as crown charter. Proprietor owns a proprietary colony, is in charge of some amount of land. 
        \item Calvert's title later goes to Lord Baltimore and Lord Baltimore II, inherit titles. They now run the area. 
        \item Pass act of toleration: Freedom to all christians, even citizenship to women and Indians.
      \end{itemize}

    \subsection{Carolinas}
      \begin{itemize}
        \item Originally united.
        \item Crown colonies given to wealthy supporters. Whole land given to twelve people. These proprietors had absolute authority over Carolinas -- Nothing like England where they were elected assemblies. Each guy had about 48,000 acres, some actually came over to colonies.
        \item Indians would sell other indians to these proprietors for material goods. They then become dependent upon white population for their goods.
        \item Eventually king divides north and south based on economics: In the south, there are large plantations for tobacco, rice, indigo, and there's market for slaves. In the North, there are smaller farms and less plantations, raising vegetables, meat, fruit, less cash crop. 80\% population white in North, almost reverse in South. 
      \end{itemize}

    \subsection{Pennsylvania}
      \begin{itemize}
        \item Swedes there first, but nobody cares.
        \item Name taken from William Penn: Left England \& Church of England to join a new protestant denomination: Quakers. Quakers had some progressive ideas, won't bear arms. Got along well with the Indians. King Charles, however, wasnt down with the Quakers, gives them Pennsylvania in order to make them leave, and so they all leave England.
      \end{itemize}
\end{document}