\documentclass{article}
\begin{document}

  \begin{center}
    {\small{} Early American History Lecture Notes} \\[0.6cm]
    {\small{} Cameron Carroll -- September 25, 2013} \\[0.6cm]
    {\small{} Lecturer: Judy Campbell, Cuyamaca College}\\[1cm]
    {\small{} Revolutionary War II}\\[1cm]
  \end{center}
  
  \tableofcontents
  \newpage
  
  \section{2nd of 3 periods of Revolutionary War (continued)}
    \subsection{Two Prong Attack}
      \begin{itemize}
         \item British enter the south with their second of two-prong attack. (The first was into New York from Long Island.)
         \item In the South, the British infect their bullets/arrows with poison. 
         \item Also during this southern second prong, Hessians, German mercenaries, were sent. They're not interested in `gentlemens war.' They beat the colonists and then killed anyone who surrendered. (Considered uncivilized atrocities, which turned some Southern loyalists to the patriot side.) 
      \end{itemize}

    \subsection{George Washington's Crossing of the Delaware}
      \begin{itemize}
        \item Washington needs a victory; Down to about 5,000 men, 6 months since Declaration, losing battle after battle and against 50,000 British and Hessians. Plus many of the 5,000 were enlisted men who were due to go home in like a week.
        \item Howe (British General) is passing through Princeton/Trenton in New Jersey. Washington comes to try to meet him to stop his progress.
        \item He decides to cross the Delaware River to sneak up on the Hessian king. So he crosses with the help of the Massachusetts Boatmen, and brings part of his regiment. He quietly gets 300 men and cover horse hooves and wheels with cloth and burlap so they can sneak up easily.
        \item Providence strikes! A dude from Washington's group just gets crazy and runs into Hessian camp. Hessian General thought he was just crazy and threw him in the brig.
        \item Providence strikes again! An informant, then, tries to warn him that Washington is close but the general is drunk and ignores it.
        \item Washington attacks in the morning, gets Hessian army to surrender, and only loses 2 guys. (They froze.)
      \end{itemize}

    \subsection{Trenton \& Princeton}
      \begin{itemize}
        \item Washington goes on to Princeton from Trenton.
        \item Realizes he can't fight traditionally; Has a decisive, guerilla victory at Princeton.
        \item In late 1770's, Washington's army is up to 8,000 people. Washington offers to pay \$10 to everybody who will stay and fight. Many stay, between this and allegiance to Washington.
      \end{itemize}

    \subsection{Generals Howe \& Burgoyne}
      \begin{itemize}
        \item English eventually do reach Philadelphia.
        \item Howe (In Philly) makes arrangements with General Burgoyne (Canada) to meet in New York, in the middle, and cut off New England from the rest of the colonies.
        \item For whatever reason, Howe doesn't go north though, and goes to try to take over Baltimore first.
        \item Because he didn't follow the plan they lose New England, goes back to Philly to regroup and wait through winter. (They had record blizzards.)
        \item Washington retreats to regroup, too, to Valley Forge, where it seriously sucks.
      \end{itemize}

    \subsection{Saratoga}
      \begin{itemize}
        \item A series of battles over several weeks, Colonists battling British.
        \item Two American generals: Gates \& Benedict Arnold. They manage to capture 5-6,000 British troops and the fort of Saratoga.
        \item Is remembered as the `turning point' of revolutionary war.
        \item Is also the point when the French enter the war. They send over 150,000 men \& 27,000 muskets. France wants revenge for the French \& Indian war. Louis XVI is helping the people to defeat the King of England, which ends up backfiring in the French revolution.
        \item Spain also enters the war.
      \end{itemize}

  \section{Final Period of the War}
    \begin{itemize}
      \item Majority of British army are taken down to Caribbean to defend those territories, breaking up the huge concentration in colonies. 
      \item About 200 Colonists manage to seize the Northwest Territory area from the British. 
      \item Around 1778-1779, morale is down despite help of French. Colonies have been fighting for 2-3 years and are tired and aren't cooperating. Congress cannot raise suport because they can't tax anybody. In order to feed/clothe the Colonial soldiers, they had to ask nearby townspeople for assistance.
    \end{itemize}
    \subsection{The Serapas}
      \begin{itemize}
        \item Americans were running hit \& run with privateers for a long time, but had built a few bigger ships. (Star of India size maybe.)
        \item In 1779, John Paul Jones; Bonhomme Richard (his ship) decides to take on Serapas (huge british warship, 50 cannons.)
        \item So he puts up British flag, sails towards Serapas, doesn't get blown away due to his disguise. So he gets alongside, puts up American flag, and starts trying to sink the Serapas.
        \item Both ships get entangled by their ropes and stuff, and start firing cannons at each other. (Through each other.) However American ship has sharpshooters keeping British from boarding their ship. Americans also start tossing grenades down the hatches.
        \item Eventually the British pull away and surrender.  But the next day the battered American ship sinks.
        \item John Paul Jones becomes father of American navy.
      \end{itemize}
    \subsection{Benedict Arnold}
      \begin{itemize}
        \item Felt like he didn't get the credit for Saratoga or the promotions that he felt he deserved.
        \item Plus he fell in love with the daughter of a British officer (half his age.)
        \item So the British offer \$100,000 and position of Brigadeer General to betray his army and so he does.
        \item Also some guy Andre gets the plans to West Point, but is caught by colonists and hung. Arnold, however, manages to get away and escape back to England. But nobody likes him in England either. 
      \end{itemize}
\end{document}
