\documentclass{article}
\begin{document}

  \begin{center}
    {\small{} Early American History Lecture Notes} \\[0.6cm]
    {\small{} Cameron Carroll -- September 30, 2013} \\[0.6cm]
    {\small{} Lecturer: Judy Campbell, Cuyamaca College}\\[1cm]
    {\small{} Revolutionary War III}\\[1cm]
  \end{center}
  
  \tableofcontents
  \newpage
  
  \section{End of the Revolutionary War}
  \begin{itemize}
    \item 1778-1779ish, 2-3 years into the war, morale is down and this war sucks. 
    \item Colonies aren't cooperating, feel dragged into the war and news of the worst blizzard in a minute was coming. Washington was holed up at Morristown and Valley Forge. 
    \item Congress is unable to get support due to our government having no way to raise money. 
    \item Benedict Arnold defects to British. 
    \item Also, the British people are tired of war as well.
    \item The campaigns in the South were the last parts of the Revolutionary War.    
    \item Southern Carolina is farming at this point and couldn't find enough people to fight without arming the slaves. 
  \end{itemize}
  \subsection{Savannah \& Charleston}
    \begin{itemize}
      \item British knew they had Southern sympathies and figured there would be an easier battle down there.
      \item They had gone in with Cornwallis \& Clinton as generals, and took over Savannah/Charleston with their navy.
      \item They moved in ~100 ships and ~9,000 soldiers, so they pretty much had complete control of the southern ports.
      \item This is the worst loss in the war for the Colonies.
    \end{itemize}
  \subsection{Patriots \& Backwoods Fighters}
    \begin{itemize}
      \item Most southern fighting is guerilla warfare. 
      \item Washington is holed up in New Jersey/Pennsylvania, and never even comes down to the South. American army hardly had shoes and would have had to walk down the entire country.
      \item They ended up relying on some patriots, good backwoods fighters...
        \begin{itemize}
          \item Nathaniel Green
          \item Daniel Morgan
          \item Lafayette (young French noble who idolized Washington, volunteered to come over and fight alongside him.)
          \item Mad Anthony Wayne (Crazy ass who helps to subdue northwest territory.)
          \item Francis Marion (Swamp Fox, pew pew pews from the swamp and then dips away real quick)
        \end{itemize}
    \end{itemize}
  \subsection{Yorktown (or Cornwallis \& Clinton Lose the War)}
    \begin{itemize}
      \item Clinton is hammered by this guerilla warfare, is pretty much over fighting in the South, and gets sent to New York, leaving Cornwallis in the South to be joined by Tarleton (a notorious guy who really hated America.)
      \item Providence keeps Tarleton from crossing a river by raising the water for 2 days after which time the Americans got away again. 
      \item Cornwallis, staying closer to coastline with his larger army, goes up through SC, NC, up to Yorktown Virginia.
      \item Coming towards him southward is Washington with his own army. Cornwallis has only 1,300 men, so sets up camp and waits for reinforcements from Clinton.
      \item Clinton, however, thinks he's being a bitch and never sends any help.
      \item Thinking he's outnumbered as usual, Washington (and Lafayette \& Von Stuben, a crazy Prussian guy) holds siege around Cornwallis instead of attacking.
      \item Cornwallis is still waiting for help, turns around looking for British help and finds only the French navy making them a sandwich.
      \item Cornwallis, here, surrenders British army ending the war.
      \item Britain, with 13,000 troops spread across colonies, lost the war with their last battle at Yorktown. People back in England were tired of the expenditures, of the war in general, and of not succeeding.
    \end{itemize}

  \section{After the War}
    \begin{itemize}
      \item This is the very first time that a man-made government creates a government without a king/emperor/whatever. Contrast this with modern attempts to overthrow the ruler in near-east countries who continue to fail to set up democracies. America, on the other hand, in the face of all these odds, manages to set up a democracy. 
      \item 7 years pass since the first shots at Lexington / Concord. It is now 1782!
      \item New York becomes temporary capital.
      \item NYC and Georgia had lots of loyalists in them. Washington \& Army give no punishment to loyalists and allow them to leave or stay as they like. About 30,000 people do emigrate to central \& South America, or Canada.
      \item International trade becomes unrestricted and booms.
      \item Treaty of Paris is negotiated in 1783 between England and the United States. This gives colonies their independence officially. United States is also given all land up until the Missisippi, doubling the size of the country.
      \item New immigration pushes west into the newly acquired land. Within first year, 4000 people move into Ohio. Ten years after, 40,000 people are in the area.
      \item American army disbands, but goes to Washington first and asks him to be king of America. George says no. (How strange!) He wants to go home instead. Congress tells the troops (some of whom had been fighting for 5-7 years) and says they'll pay them \$80 for their service... but as an I-O-U. 
      \item Slavery is abolished in Pennsylvania northwards. 
      \item Soldiers go home... with some riots even, in Pennsylvania... disgruntled at being not directly paid.
      \item Currency at the time was a Continental. Government  (Continental Congress) decided they would just print some money and distribute it. This causes inflation. 
      \item Why did we win? Arguably it was because Washington kicked ass; Without his leadership and hope America probably would have lost. Also because the French: their navy, ammunition, and help were also instrumental in winning the war. 
    \end{itemize}

  \section{Northwest Territories}
    \begin{itemize}
      \item Articles of Confederation saw us through the war, too: Made peace treaty and also made Northwest Territory. 
      \item Northwest Ordinance of 1787 established how we get new states: Western land would be occupied and settled. As they become settled, they become territories first. Territories don't pay taxes but also aren't represented by government.
      \item First state after the thirteen is Ohio.
      \item Thomas Jefferson, surveyor, goes to Ohio, and sectioned the state off into a grid. Each gridbox was given land appropriated for a school. Jefferson thought everyone should be able to read \& write.
      \item No slavery allowed in Northwest Territories
      \item All states given same rights as one of original 13 colonies.
      \item Indians allowed to stay unless they cause problems. (Supposed to be given rights here.)
    \end{itemize}
\end{document}