\documentclass{article}
\begin{document}

  \begin{center}
    {\small{} Early American History Lecture Notes} \\[0.6cm]
    {\small{} Cameron Carroll -- October 14, 2013} \\[0.6cm]
    {\small{} Lecturer: Judy Campbell, Cuyamaca College}\\[1cm]
    {\small{} Washington, Adams \& Jefferson}\\[1cm]
  \end{center}
  
  \tableofcontents
  \newpage
  
  \section{Washington's Presidency}
    \subsection{Whiskey Rebellion}
      \begin{itemize}
        \item Pennsylvania farmers didn't want to pay a tax on whiskey because they manufactured it.
        \item Under the old Articles of Confederation, the government had no teeth, so no one was apt to pay up.
        \item Those who didn't want to pay rebelled, and Washington sent in 15,000 soldiers to keep the law. This sends the message that all inhabitants will obey the laws of the land, establishing the governents `teeth.'
      \end{itemize}
    \subsection{Hamilton's Economic Program}
      \begin{itemize}
        \item Financial program had to be undertaken. Alexander Hamilton was in charge of setting up an economic system.
        \item He sets up the national bank with a 20-year charter. (After 20 years the bank and system will be reviewed to see if they're effective.)
        \item He also sets up the US Mint. (In Washington D.C.)
        \item Also wants to fund national debt at face value. (All states' debt will be take over by the government, which will pay them back.)
        \item Wanted to, and succeded in getting the county on firm financial/economic footing.
        \item Government passes some taxes and tariffs.
        \item Thomas Jefferson wasn't down with the financial bank \& system, arguing that the people in charge could get out of control. Was aware of the problems it could (and later does) produce. 
      \end{itemize}
    \subsection{John Jay \& Naval Security}
      \begin{itemize}
        \item Internationally at this time, France is at war in Europe. Washington is trying to stay neutral, but this is sort of difficult.
        \item British would `impress' American sailors \& ships, which is British boarding the ships and taking them to sail for Britain instead.
        \item Washington sends over John Jay, who negotiates and buys peace with the British for a time.
        \item John Jay was burned in effigy up \& down the colonies because they wanted the British punished, and not just made peace with.
        \item France now starts impressing American sailors; Then Spain gets involved as well. Washington just wants to trade with everyone, but this becomes increasingly difficult.
      \end{itemize}
    \subsection{Death of Washington \& the Birth of Political Parties}
      \begin{itemize}
        \item Around this time, Washington dies. (67 years old at Mt Vernon, died from leeches trying to cure his throat virus.)
        \item After Washington dies, political parties start to form. (Washington was sorta holding everyone together before.) 
        \item Federalists (including Washington) were pro-England. They argued that their heritage was British, we're basically Englishmen and we should support Britain.
        \item The Democrat/Republicans, on the other hand, were pro-France, arguing that they helped us get our independence. 
      \end{itemize}

  \section{John Adams' presidency}
    \begin{itemize}
      \item Problem surfaces with John Adams: Whoever got most votes was President, second runner-up was Vice, according to the pure popular vote. John Adams won, and Jefferson was second-runner-up; The former was an Antifederalist and the latter a Federalist, with very different ideas about how the government should be run.
      \item Adams was the first to live in the Executive Mansion! At this time, however, DC was a swampland. The Presidential Mansion had only 5 usable rooms, 13 fires to keep the place warm, basically not super luxurious.
      \item Adams' cabinet was more loyal to Jefferson than to him. 
      \item Only in office for 4 years.
    \end{itemize}
    \subsection{The XYZ Affair}
        \begin{itemize}
           \item French didn't like him because he didn't support them.
           \item They start harassing our shipping. Adams sends a delegation to France, the XYZ affair, to say `stop it you guys.'
           \item The French try to bribe the American ambassadors. Adams is quoted as saying `millions in defense, but not a cent in tribute.' Of course, over the years we've paid tributes but... you know... whatever.
           \item This made Adams very popular because he stood up to the French.
        \end{itemize}
      \subsection{Alien \& Sedition Acts}
        \begin{itemize}
          \item Tries to weaken with Democrat/Republican party with the Alien \& Sedition acts.
          \item Wanted to raise the hurdle for citizenship. At this time, people came into the country legally but they didn't become citizens automatically.
          \item Federalists want a high hurdle for becoming a citizen, while Democrat/Republicans who catered to poor and/or uneducated masses and wanted a lower hurdle.
          \item Sedition Act: Tried to muzzle the newspapers by making it a crime to say anything bad about the President.
          \item Neither act passes. 
        \end{itemize}
      \subsection{Kentucky \& Virginia Resolutions}
        \begin{itemize}
          \item If a state thinks the federal goverment is acting unconstitutionally, the states have a right to disregard it. 
        \end{itemize}
      \subsection{John Marshall}
        \begin{itemize}
          \item Adams appoints a Supreme Court justice, John Marshall (his cousin.)
          \item The president alone nominates a supreme court justice although congress has to `okay' them.
          \item Supreme court justices far outlast the president that elected them.
          \item Marshall is known as the most influential justice that we've had.
        \end{itemize}

    \section{Jefferson's Presidency}
    \begin{itemize}
      \item Wrote the Dec.
      \item Seen as a `flaming' liberal, even though he wouldn't be wild by today's standards.
      \item Sort of a renaissance man, although a poor public speaker. 
      \item Served in Virginia House of Burgesses.
      \item Secretary of State (to Washington?)
      \item Fun Fact: His grandson was the first baby born in the White House!
      \item When Jefferson was elected, it was sort of remarkable because our country switched from Federalist to Republican/Democrat. 
      \item For Jefferson's election (in 1800, now using Electoral system), Democrat/Republicans won the popular vote but tie in terms of Electoral votes. The vote goes to Congress, and Hamilton shifts the vote toward Jefferson. Now we're voting for a President and also Vice: They're running for a specific office, rather than for whatever they can get. 
      \item Chooses his cabinet from people he thinks will be most competent; Madison is Secretary of State; Galacian for Treasury.
      \item Jefferson wanted an agrarian nation; He had seen Paris \& London: soot-filled cities of the industrial revolution and wanted something different for the U.S.
      \item He supported education strongly.
      \item Idea of military, like Democrat/Republicans, bordered on pacifism. Pulled back on support of the military.
    \end{itemize}
    \subsection{Louisiana Purchase}
      \begin{itemize}
        \item Jefferson wanted access to the Missisippi river.
        \item At this time, Napolean is wreaking havok (Napoleonic wars) and running out of money.
        \item The opportunity presents itself, then, for Jefferson to purchase all the land in the Louisiana Purchase. 
      \end{itemize}

\end{document}