\documentclass{article}
\begin{document}

  \begin{center}
    {\small{} Early American History Lecture Notes} \\[0.6cm]
    {\small{} Cameron Carroll -- October 16, 2013} \\[0.6cm]
    {\small{} Lecturer: Judy Campbell, Cuyamaca College}\\[1cm]
    {\small{} Jefferson, Madison \& Monroe}\\[1cm]
  \end{center}
  
  \tableofcontents
  \newpage

  \section{Jefferson's Presidency}

    \subsection{Marbury vs Madison}
      \begin{itemize}
        \item Was an important court case during Jefferson's presidency. 
        \item This establishes judicial review, ability of supreme court to determine what is constitutional. (One of the ways constitution keeps current.) (John Marshall was supreme guy.)
        \item Adams appoints Marbury, a federalist, before he goes out of office. Adams writes the letter, but it doesn't get delivered to Marbury until later. Then Jefferson becomes president and says to not deliver the letter because he's a Republican. Marbury, then, sues Madison because he wants to be a judge. 
        \item This case goes to supreme court (John Marshall) which throws Marbury under the bus. He tells Jefferson that delivering the appointment would be the `right thing to do' but doesn't force him to do it.
        \item Marshall knew he couldn't fight Jefferson; He had too much support.
        \item He sets a precedent for judicial review, deciding that it was constitutional to \emph{not} deliver the letter.
      \end{itemize}

    \subsection{Louisiana Purchase}
      \begin{itemize}
        \item Monroe, Livingston sent to France to ask Napoleon if they could buy port of New Orleans. Napolean needs money, so he's pretty down. He actually offers to sell the entire Louisiana area for \$15 million dollars, including the port of New Orleans, and they accept.
        \item This doubles the size of the United States. Jefferson is stoked, but constitutionally the President can't just buy land and call it the United States. 
        \item `Essex Junto:' (New England area says the president acted illegally and unconstitutionally and argued that they have a right to withdraw.) But I mean come on it was a great deal, so people get over it. But the thought that `if we don\'t like what's going on we can just leave' remains.
        \item Jefferson sends Louis \& Clark out to explore what he just bought.
      \end{itemize}

    \subsection{Burr Conspiracy}
      \begin{itemize}
        \item Burr was vice president during Jefferson's term, but nobody liked him (even Jefferson.) For the only time in US history, he chooses a different vice president for his second term.
        \item Burr then tries to become governor of New York. But he loses because Hamilton, a well-known dude, campaigns for the other guy.
        \item Burr gets pissed off, challenges Hamilton to a duel. But Hamilton's son had died in a duel, and so he just fires his gun in the air and gets shot dead.
        \item Everybody is even more pissed off at Burr, so he dips out to New Orleans with a group of military dudes. He crosses into French/Mexican/Spanish territory to Texas to try to form his own country. US Army captures him and brings him to trial.
        \item This goes all the way to the supreme court but John Marshall acquits him based on a duel being a legal way to settle differences. 
      \end{itemize}

      \subsection{Barbary Wars}
        \begin{itemize}
          \item In 1801, Jefferson sends what navy we have to the Mediterranian.There were pirates in the area attacking American trading ships, asking for ransom.
          \item Unfortunately Jefferson downsized the navy. But it does kinda succeed in stopping the pirating.
          \item He puts in an embargo act, but this hurt Americans and our economy more than it hurt the pirates.
        \end{itemize}

      \subsection{Chesapeake Leopard Affair}
        \begin{itemize}
          \item Britain backed on on impressment, but one British ship goes into the Chesapeake bay, stops an American ship and impresses them on the stop saying they were escaped Englishmen.
          \item Jefferson was hard-pressed to stay out of war, felt that way himself. Leaves it for the next dude.
        \end{itemize}

        \paragraph{} Jefferson never even mentions being President on his epitaph, and voluntarily skips a third term.


  \section{James Madison}
    \begin{itemize}
      \item Was a tiny dude, short and only 100 pounds.
      \item His wife was 17 years younger, very fancy, instrumental in getting those people who \emph{needed} to talk to her husband at social events to him, and keeping others away.
      \item Known as Father of the American Constitution.
      \item Wrote 9/10 amendments for Bill of Rights.
    \end{itemize}

    \subsection{War of 1812, Mr. Madison's War}
      \begin{itemize}
        \item Also called Second War for Independence. A show that nobody's gonna take us over again.
        \item Fought over freedom of the seas; Britain and France were both impressing sailors. Also wanted to be free to trade.
        \item Most fighting around Great Lakes area.
        \item British in Canada were arming the Indians. Indians were uprising, causing havok among frontiersmen.
        \item Since Revolutionary War, lots of people in Congress felt that we should have taken Canada when we got independence. (War Hawks)
        \item Spain still sorta-owns Florida, and we figure we may as well take that also. (Or the War Hawks do at least.)
        \item United States is taking an expansionist sort of view.
        \item At this time, Jefferson let the national bank charter expire after its 20-year limit. He was a Republican, and didn't believe in the central bank. 
        \item Madison can only get like 5,000 soldiers because of 
        \item All of Britain's forces were not concentrated in this area, they were busy with other things. (Napoleon)
        \item Most important part of war was when Napoleon is finally defeated, and England can finally deal with these pesky United States, and she sends entire army and navy to the US. Moves up and down the coast, burning cities due to their overwhelming strength. \\
        Also goes into Washington D.C. to burn it as well. \\
        In August, when they reach D.C. it's clear that they're coming up the river. The President and Congress all dip out, but Dolly stays behind to save all the artifacts they can. They cut canvases out of the frames, roll them up and take them. The portrait on the \$1 bill is one of the ones she saved. They barely escape the British. \\
        British burn the Presidential Mansion to the ground.
        \item British then go up the Chesapeake to burn down Baltimore. On the way to Baltimore, though, is Fort McHenry, they have a big fight and some guy writes our national anthem about what happened.
        \item British attacked DC and New Orleans both. Dude Andrew Jackson is from Tennessee, where Indians were on a rampage and had massacred 250 frontiersmen. \\ 
        Decides this won't go unpunished... takes off into Alabama after them, where he massacres like 900 Indians viciously. \\
        Jackson heard about the British going to New Orleans, decides to head down there. He collects the `long-rifles' from frontiersmen, pirates, escaped slaves from Haiti, and ends up with like 5400 of rabble, fights in New Orleans. \\
        They get there before the British do, and get up on the hills. British find out they have to scale the hills but have to take two weeks to make enough ladders to assault Jackson's position. \\
        A fog comes in, British guys think it's great cause Jackson won't see them coming, but they're wrong cause when they get up on the hills Jackson kills 2000 British and loses only 8.
        \item The British withdraw from the Chesapeake, sign a peace treaty, and we've never fought since.
      \end{itemize}

    \subsection*{}
      \begin{itemize}
        \item Madison was credited with getting us into the war, but he did actually succeed in getting us out.
        \item He re-charters the National Bank.
        \item Dude is featured on the \$5,000 bill. 
      \end{itemize}

  \section{James Monroe's Presidency}
    \begin{itemize}
      \item Last of revolutionaries to be president.
      \item Had `era of good-feelings' under this guy. (1820's, 1830's)
      \item Was a blending of political ideologies. 
      \item Sends Andrew Jackson down to Florida, makes it part of the United States.
    \end{itemize}

    \subsection{Monroe Doctrine}
      \begin{itemize}
        \item In years following our own Revolutionary war, we had a French revolution and many other countries realized that their holding country wasn't as strong as they thought.
        \item Monroe thought that this wasn't a good idea economically for these countries to be so unstable. Their revolutions weren't anywhere near as organized as our own (which was a hot mess.)
        \item Monroe Doctrine is issued to the rest of the world, saying three things.
        \begin{itemize}
          \item No new colonies in the Western Hemisphere!
          \item If any colony wants its independence, Europe must yield it with no war.
          \item As a result, he says that the US will not interfere in foreign affairs outside of the Western Hemisphere. (Still holds today.)
        \end{itemize}
        \item There since has been a caveat added to the third point: `...unless asked.'
      \end{itemize}

    \subsection{Missouri Compromise}
    \begin{itemize}
      \item First state to enter the Union was Missouri.
      \item People were moving into this area from slave states.
      \item In Congress at the time there was a split between slave states and free states.
      \item If Missouri was admitted as a slave state OR a free state, that would throw off the balance in Congress.
      \item Henry Clay, the Great Compromiser, says we should bring Maine in as a free state and Missouri as a slave state and keep balance. (So they do.)
      \item Missouri Compromise:
      \begin{itemize}
        \item Missouri comes in as slave state
        \item Maine comes in as a free state
        \item At 36rd Parallel, a line was drawn. Anything north of that line must come in as a free state, and anything south of the line \emph{could} come in as a slave state. 
      \end{itemize}
      \item Southerners at the time were so used to free labor, their economy basically depended on it.
    \end{itemize}
\end{document}