\documentclass{article}
\begin{document}

  \begin{center}
    {\small{} Early American History Lecture Notes} \\[0.6cm]
    {\small{} Cameron Carroll -- October 21, 2013} \\[0.6cm]
    {\small{} Lecturer: Judy Campbell, Cuyamaca College}\\[1cm]
    {\small{} John Quincy Adams \& Andrew Jackson}\\[1cm]
  \end{center}
  
  \tableofcontents
  \newpage

  \section{John Quincy Adams' Presidency}
    \subsection{Fun Facts}
      \begin{itemize}
        \item Only son of a president to become president himself, up until the Bushes.
        \item Born 8 years before shots at Lexington \& Concord. 
        \item Many of the trees planted at white house were done by him.
        \item Didn't particularly like being President.
        \item Known for starting the Smithsonian.
        \item Was the only President to go back to being a Congressman, liking that better. He died of a heart attack on the floor at 80 years old.
      \end{itemize}

  \section{Andrew Jackson's Presidency}

    \subsection{Age of Jackson; Jacksonian Democracy}
      \begin{itemize}
        \item Age of Jackson centered around his personality, which was both good and bad.
        \item Was an army man. At this time, presidential candidates could either be army guys or lawyers.
        \item Was from a frontier immigrant family, first president from non-mover-shaker-type family.
        \item \textbf{Old Hickory:} Liked to play up the idea that he was just a common guy, even though he was a wealthy plantation owner when he was elected.
        \item Is a tall, handsome, popular guy. Dresses sharply, doesn't look like the common farmer guy.
        \item Was in two duels, killed a guy for insulting his wife. Known for a volatile temper.
        \item While in office, he had to had bullets removed from his arms and walked around with another unrecoverably embedded inside him.
        \item Fought in Revolutionary war at a very young age. Hated the British.
        \item First president to have an assassination attempt on his life. Only survives by some serious providence (two gun misfires.) Jackson tackles the guy and beats him with his cane. Later, nothing is found to be wrong with the guns. They just didn't work on Jackson. First case of a plea of insanity holds up in court.
        \item Became a national hero after Battle of New Orleans.
        \item Beats up on the Seminole Indians in Florida.
        \item After Monroe's presidency there was political blending and the `era of good feelings.' But now we have the two political parties returning...
          \begin{itemize}
            \item Federalists \& Whigs $\rightarrow$ Republican Party
            \item Anti-Federalists $\rightarrow$ Democrat Party
          \end{itemize}
        \item Age shows the first signs of campaigning, and the start of the national political conventions.
        \item Jackson is really well known for his Indian policy. He sided with states rights over federal government decisions. He hated Indians despite having adopted an Indian boy.
        \item Like Washington, Jackson makes the executive branch very strong. But Washington did it in a fair manner, but Jackson does it with force. (And gets away with it because of his personality.) He uses the veto more than anyone else in history.
        \item Nation does become more democratic under him; There's a strong focus on the common man.
      \end{itemize}

      \subsection{Thomas Mast}
        \begin{itemize}
          \item Created the American version of Santa Claus.
          \item Gave Republicans the symbol of an elephant:
            \begin{itemize}
              \item Democrats find it obtuse.
              \item Republicans find it dignified and strong.
            \end{itemize}
          \item Gave Democrats the symbol of a donkey:
            \begin{itemize}
              \item Democrats find it humble and smart.
              \item Republicans find it stubborn and a bit of a jackass.
            \end{itemize}
        \end{itemize}

      \subsection{Webster Hayne Debate \& Sectionalism}
        \begin{itemize}
          \item Debate over how to settle all the land just acquired.
          \item Westerners, people in this area wanted the government to just give them the land so they can farm it and occupy it, but they can't pay a lot of money.
          \item Southerners wanted this land for free or cheap, also, needing to expand with their cotton and tobacco industries.
          \item New England states, however, think the land has to be sold in order to bring money into the treasury. They're considering that all the immigrants coming into the country come through the ports particularly in the New England area. They were reliant upon this cheap, essentially captive labor force and didn't want them making an exodus to the West.
          \item States were ready to separate again, all had different sectional issues.
          \item Congress decided that the land isn't divisible and so we have to agree on compromises. Pushed for the union of the United States. 
          \item Land was to be sold fairly cheap, but not given away.
        \end{itemize}

      \subsection{Tariff of 1832}
        \begin{itemize}
          \item Jackson was in support; Federal Government wanted to tax manufactured goods coming into the country. (Buy American!)
          \item Northern States, of course, are all for it as the manufacturers.
          \item Southern States \& Western States which needed to buy manufactured goods, wanted the best possible price.
          \item Calhoon from South Carolina said that their state refuses to pay the tax. Jackson says they will and tries to send the army down to lynch Calhoon. Henry Clay, known as the Great Compromiser, says to the president that he can't just lynch people though. 
          \item As a compromise, they lower the tax and this seems to hold the states together (for a while.)
        \end{itemize}

      \subsection{Dissolution of National Bank}
        \begin{itemize}
          \item The second 20-year charter period of the National bank comes up again.
          \item Bank run by Biddle, who was very corrupt. He made commercial loans to friends \& manipulated interest rates. Jackson saw this and combined with his own ideologies, he decided not to renew.
          \item Instead, he decides to dismantle the Second Bank of the United States and allows each state to charter their own.
          \item Now there is no longer a central currency, allowing each state to mint their own.
          \item When Washington was President there were 24 charter banks, but while Jackson is president the count goes up to about 700.
          \item `Pet Banks' were created, with very little supervision. Half of them were shut down with no federal backing.
          \item This leads to inflation and to combat the inflation, Jackson enacts specie circular, which says that there must be gold and silver backing to the currency used for land purchases. This caused an increase in the value of the land.
        \end{itemize}

      \subsection{Indian Removal Act}
        \begin{itemize}
          \item In Georgia \& Alabama, Cherokee owned a huge reservation given to them by the US government. 
          \item This reservation operated according to US government rules, but wasn't officially part of the US.
          \item The society was literate, had  a government and law system based on the US, but operated as a separate entity. 
          \item Unfortunately, gold was found in the Appalachian mountains on their reservation. The people around them wanted the ability to go mine for the gold. The Cherokees said no thats not cool.
          \item This issue goes to the supreme court who rule that nobody can just go onto the territory and mine, saying it belongs to the Indians.
          \item Jackson defies the supreme court, sends the troops down to Georgia to move the Indians off of the land.
          \item He forces them to walk to Oklahoma, says they can have that area and sent the army with them to make sure they got there.
          \item Known as the Trail of Tears because so many people died along the way and they couldn't follow their custom of having a burial ground.
          \item Winfield Scott is the general in charge of moving all the Indians.
        \end{itemize}
\end{document}