\documentclass{article}
\begin{document}

\begin{center}
  {\small{} Early American History Lecture Notes} \\[0.6cm]
  {\small{} Cameron Carroll -- November 13, 2013} \\[0.6cm]
  {\small{} Lecturer: Judy Campbell, Cuyamaca College}\\[1cm]
  {\small{} Sectionalism \& Starting the Civil War}\\[1cm]
\end{center}

\tableofcontents
\newpage


\begin{itemize}
  \item Southerners had an `air of superiority' about them; They saw the South as an area of prosperous planters where the slaves were treated well. However, most people in the South were illiterate and lived on the frontier as small, independent farmers. Most whites in the South owned 4 slaves or fewer.
  \item Slave codes said that no slaves could own guns or leave the property without permission, and also were not supposed to be educated.
  \item Despite being slaves, their owners had made an investment... They take care of their slaves by feeding, clothing, housing them.
\end{itemize}

\section{Justifications for Slavery}
  \begin{itemize}
  \item Scientific racism: The idea that blacks were inferior physically and psychologically, and people legitimately believed this. 
  \item Historic: All great empires have had slaves in their societies.
  \item Religious: There's slavery in the bible. People were taken out of Africa, sprinkled with some holy water, made Christian and sent to be slaves. 
  \item It was also argued that it was better for them to be working as a slave as compared to the factory workers in New England. 
  \end{itemize}

\section{Wilmont Proviso}
  \begin{itemize}
    \item The question arose... `What do we do with states coming into the union? How do we balance free and slave states?'
    \item At this time there were no countries in the world that legally allowed slavery.
    \item Currently, At the $36\,^{\circ}$ 30\'\ line, anything north must be free but anything south could come in as a slave state.
    \item California wants to come in as a free state but there's no slave state to balance it.
    \item Southern states were feeling rather gipped. Right after the Mexican Cession, Wilmont Proviso proposed there will be no slaves at all in the Cession area.
    \item Congress doesn't pass this proposal.
  \end{itemize}

\section{Compromise of 1850}
  \begin{itemize}
    \item California is allowed to enter union as a free state but the Utah and New Mexico territories are brought in under popular soverignty, allowing them to decide for themselves between free and slave states.
    \item Before 1850, the Fugitive Slave law which said slaves are property and therefore should be returned to their owner's state wasn't really enforced. After 1850, however, it is enforced more heavily.
    \item In Washington DC, because it looked tacky for slave markets to be on the streets in the capital, there was no slave trade allowed on the streets.
  \end{itemize}

\section{Kansas-Nebraska Act of 1854}
  \begin{itemize}
    \item At this time, Kansas and Nebraska are territories.
    \item This act allows both to come in as states under popular soverignty as well.
    \item However by doing this, Congress negated the Missouri Compromise and the 36' 30'' line.
    \item This causes a mess in Kansas, which becomes known as Bleeding Kansas:
      \begin{itemize}
        \item The Northern Abolitionists decide that they should go populate the area if it's going to be popular soverignty.
        \item However Southerners move into the area as well.
        \item People begin raiding each others farms and killing whomever disagreed with them.
        \item Oklahoma at the time is still Indian territory, who by the way own slaves.
      \end{itemize}
  \end{itemize}

\section{Movers \& Shakers Before the War}
  \begin{itemize}
    \item Nat Turner takes 50 guys, free and slave blacks, and kills a bunch of white slaveowners. They're caught and killed. However the southerners get goosey about the possibilty of slaves organizing a revolt.
    \item William Lloyd Garrison starts a newpaper, the Liberator, and was a very strong abolitionist speaker.
    \item Harriet Stowe, another abolitionist, write Uncle Tom's Cabin.
    \item In Congress:
      \begin{itemize}
        \item Calhoon
        \item Jefferson Davis (Later becomes President of the South)
        \item Steven Douglas (Ran against Lincoln)
        \item Cassius Clay (Emancipationist, minister to Russia)
        \item Seaward (Buys Alaska as Secretary of State to Lincoln)
      \end{itemize}
      Abolitionists wanted to be like the rest of the world, but Southerners argued that the federal government was trying to take private property.
      \item A fight erupts on the floors of congress between Sumner and Brooks. Sumner accuses Brooks of taking the harlot of slavery. Brooks beats the utter crap out of Sumner in return.
      \item Dred Scott:
        \begin{itemize}
          \item Was a slave whose owner moved to Illinois, a free state.
          \item He becomes more of a servant since slavery is outlawed. But the owner wants to move back to Missouri and Scott doesn't wanna go back to being a slave.
          \item This goes all the way to the supreme court who rule `once a slave, always a slave.'
          \item The Dred Scott Case ended the prohibition of slavery in federal territories and prohibited Congress from regulating slavery anywhere, overturning the Missouri compromise, enabling popular sovereignty, and bloody Kansas.
        \end{itemize}
      \item John Brown was a white abolitionist who led a rebellion of other abolitionists and raided an army arsenal at Harper's Ferry. He was surrounded and caught and hung for treason, but becomes something of a martyr.
  \end{itemize}

\section{Election of 1860}
  \begin{itemize}
    \item Republican Lincoln ran against Democrat Steven Douglas. (There was also the Union party.)
    \item Lincoln was running as the poor, working man's party and thought slavery was wrong, but not quite an abolitionist.
    \item In October of 1860, South Carolina says that if Lincoln wins they will secede.
    \item Sure enough, Lincoln is elected with only 40\% of the vote. 
    \item With Lincoln's victory, seven states making up the Lower South secede.
  \end{itemize}

\section{Formation of the Confederate States}
  \begin{itemize}
    \item Confederate states form with Davis as their president and Stevens as his vice. They were set up basically just like the United States.
    \item They seceded because they just thought division was a reasonable thing and that the North would just let them go. They knew politically they probably couldn't keep slavery forever.
    \item Around this time, West Virginia splits from Virginia. (It's on the other side of the Appalachians.) They decide to try to stay neutral with their own little government.
    \item Virginia, North Carolina, Tennessee and Arkansas didn't like beign called traitors but just felt that slaves were necessary to their economy and way of life.
    \item This results in border states, those that didn't believe in secession. (Missouri, Kentucky, West Virginia, Maryland, Delaware) The border states had more people than all the rest of the south, and half of the factories. Plus, Maryland basically has our capital.
  \end{itemize}

\section{Starting the War}
  \begin{itemize}
    \item In March 1861, Lincoln sneaks into DC a few days early, afraid of an assassination attempt.
    \item In Fort Sumter in South Carolina, the war begins:
      \begin{itemize}
        \item The South had taken over all the forts along the coast leaving the union only two.
        \item Sumter was held by Major Anderson (Union). There was no reason to stock up the forts with a whole bunch of guys, and there are only like 75 guys there.
        \item Anderson asks for supplies from Lincoln who agrees and lets the South know it's just supplies and not military stuff. 
        \item The South says no that's an act of war and so they fire upon Fort Sumter, and amazingly only a horse was killed.
        \item But  When Southerners fired on Northerners, the Union was provoked to fight. (Americans don't shoot each other! Come on!)
        \item Up until now, the North was thinking maybe they could let them go but now called them traitors. 
        \item North asked for volunteers and whole towns went to sign up to fight.
        \item In order to keep border states, Lincoln said that the cause of the war was to preserve the union and not to end slavery.
      \end{itemize}
    \end{itemize}

\end{document}