\documentclass{article}
\begin{document}
  \begin{itemize}
    \item Cameron Kalledsoe Carroll, Kalledsoe being my mother's middle name given to my brother and I to try to carry on that part of the family's name. This is a little disappointing since the two main familial arteries are the Kornerubs and the Helms, both of which have bound volumes of the genelogical research done back in Denmark. The former can trace its lineage back to Ole Worm, a Danish physician and scholar. Worm, aside from having some more significant and mundane contributions to the world, was working on some truly great research... He finally disproved the existence of unicorns, since everybody had seen \emph{plenty} of narwhals. He also established that lemmings were really just rodents and didn't simply materialize out of thin air. 
    \item It's pretty interesting what you can dig out of the internet. You might find out you're related to someone slightly famous, or find your relatives in a newspaper from the 1800's.
    \item But this is supposed to be more about me and less about Ole Worm...
    \item I'm gonna be a little bit pessimistic for a minute... It's been said that Grossmont `is where dreams go to die.' I mean, sure, I've seen some really smart people transfer away within two or three semesters, but I've also seen it live up to this reputation. I don't think I'm about to cross the event horizon, but I still feel the pull of the black hole -- I've been here so long and have so many units that without a petition I'll be in the very last category to sign up for classes. And then, surely, the attraction will be that much stronger.
    \item I'm a Mechanical Engineering student who, as mentioned, has been here a minute. It literally was a different person who walked in the doors in 2008 to fumble through his classes, and thanks to him (and to a lesser extent, myself) I barely remember what happened. Something about wine babies on the beach I think.
    \item As for the future, I'd just like to transfer to SDSU for right now. For a long time I had a dream of going to UCSD for their Nanoengineering program, but between wanting a generalist education and not having the grades, I settled for Mechanical.

  \end{itemize}

  \begin{itemize}
    \item Good afternoon everybody. My name is Cameron... Cameron Kalledsoe Carroll. One of the two things I actually got to raise my hand for during our audience analysis was to say that I had traveled to Denmark with my mom and grandma, but I just said we were traveling... I don't think I communicated that they are both very, very Danish. 
    \item Being very Danish, my mom went ahead and gave me her father's surname as a middle-name, leaving me to wonder while growing up why I couldn't just be Cameron James Carroll or something normal.
    \item Now, though, I don't lament having the weird Danish middle maiden surname so much as I lament having Kalledsoe instead of Helm or Kornerub, the two \emph{big} familial arteries. Both of these heritages come with bound, printed books on their ancestry describing most of the people, what they did, when they died, et cetera.
    \item One of them, the Korenrubs I believe, can trace the lineage back to Ole Worm, a Danish physician and scholar. This guy, aside from having some real and significant contributions to the world, was working on such world-changing research such as finally disproving the existence of Unicorns. (The horns were obviously from Narwhals.) Despite this research, he somehow still thought perhaps Unicorn horns could possibly cure the poison he proceeded to administer to some animals. In another tour de force, he also established that lemmings were rodents and \emph{not}, as commonly believed, simply materialized out of thin air.
    \item Continuing this tradition of important research, I am apparently finding out just how much of your life you can spend at Grossmont. It's been said (a few times) that `Grossmont is where dreams go to die.' -- There is some truth and some hyperbole to this statement, but I can't deny that I feel the pull of the black hole sometimes. I'm hardly on the tip of the event horizon... transfering to SDSU is well within reach, but I know that life is a lot harder for people in other, nontechnical majors: I had an engineering class start with 8 people, and I've seen English classes get canceled with 20.
    \item It was literally a different person who first walked in the doors in 2008. But, after something like 8 semesters and changing my mind almost as many times, I'm finally looking one or two semesters until I can leave. And I cannot possibly communicate how exciting that is.
  \end{itemize}
  
.
  \begin{itemize}
    \item Ole Worm: http://jamesgray2.wordpress.com/2013/05/02/ole-worm-and-thomas-bartholin/ 
  \end{itemize}
  

\end{document}